\section{Introduction}

\subsection{Overview}

Consider the problem of determining whether a given multivariate polynomial $P$ with integer coefficients has integer roots. While this may seem simple, this problem is in fact \emph{undecidable}. That is, one cannot write a program on a computer that correctly outputs the answer `yes' or `no' to the problem (for any polynomial $P$). The focus of this course is to study such fundamental limits to computers and computation.

Let us look at the problem of determining whether a certain word is present in a `language'. For example, consider
\[ L_1 = \{ a^n b^m : n,m\ge 0 \}. \]
Now, we must write a program that given a string $s$ over the alphabet $\{a,b\}$, determines if $s \in L$. This program is in the form of a \emph{discrete finite automaton}.

\begin{center}
    \begin{tikzpicture}[node distance=3cm]
    \node[initial, accepting, state] (1) {S};
    \node[state] (2) [below right of=1] {R};
    \node[state, accepting] (3) [above right of=2] {A};

    \draw (1) edge[loop above] node[align=center]{$a$} (1)
    	  (1) edge[above] node[align=center]{$b$} (3)
    	  (3) edge[loop above] node[align=center]{$b$} (3)
    	  (3) edge[right] node[align=center]{$a$} (2)
    	  (3) edge[loop below] node[align=center]{$a,b$} (3);
    \end{tikzpicture}
\end{center}

How would we do this if we instead of have the language
\[ L_2 = \{ a^n b^n : n \ge 0 \}? \]
It may be shown that such a language cannot be recognized by a deterministic finite automaton. To create an automaton that does recognize it, we require an additional \emph{stack}, which gives rise to the \emph{pushdown automaton}. The automaton itself is almost identical to the above automaton, except that we `push' an $a$ onto the stack when we read an $a$, and `pop' an $a$ when we read a $b$. Finally, we further require that the stack is empty when the string has been read.\\

We also have what is known as a \emph{context-free grammar}. We have a bunch of rules, and generate strings in the language by repetitively performing a replacement using one of the rules. It turns out that these are equivalent to pushdown automata.\\

Finally, consider the language
\[ L_3 = \{a^nb^nc^n : n \ge 0\}. \]
This cannot be recognized by even a pushdown automaton. This leads to the \emph{Turing machine}, where instead of a stack we have a `tape'. This represents the `ultimate' computer that can do anything a computer can do. We shall look at each of these in detail over the next few sections.\\

The focus of our study shall be each of the following.

\begin{center}
\begin{tabular}{|c|c|}
	\hline Machine & Language \\
	\hline
	Discrete finite automaton (DFA/FSA) & Regular expressions \\
	Pushdown automaton (PDA) & Context-free grammars \\
	Turing machine (TM) & Unrestricted grammars \\
	\hline
\end{tabular}
\end{center}

We can further introduce non-determinism in each of the three automata. We shall see that the expressive power of DFAs and TMs do not change on allowing non-determinism, while that of PDAs does.\\

First, before explaining anything, let us set up some notation and definitions for the rest of this course.

\begin{fdef}
	An \emph{alphabet} $\Sigma$ is a non-empty set. Its elements are referred to as \emph{letters} or \emph{terminals}.\\
	The set $\Sigma^*$ is the set of all finite strings over $\Sigma$. In particular, $\Sigma^*$ contains the empty string denoted $\epsilon$. The set $\Sigma^+$ is equal to $\Sigma^* \setminus \{\epsilon\}$.\\
	A \emph{language} is a subset of $\Sigma^*$.
\end{fdef}

The \emph{Chomsky hierarchy} represents the increasing complexity of languages. At the lowest level, we have \emph{regular languages} that are recognized by \emph{finite state automata}. This is a subset of \emph{context-free grammars}, which are recognized by \emph{pushdown automata}. This in turn is a subset of \emph{unrestricted grammars}, which are recognized by \emph{Turing machines}.\\

A finite state automaton is a tuple $(Q,\Sigma,\delta,q_0,F)$, where $Q$ is a finite non-empty set of states, $\Sigma$ is an alphabet, $\delta : Q \times \delta \to Q$ is the transition function, $q_0 \in Q$ is the initial state, and $F \subseteq Q$ is the set of accepting states.\\

Next, let us consider the language $\{a^nb^n : n \ge 0\}$ we mentioned earlier. This can be recognized by the context-free grammar
\begin{align*}
	S &\to \epsilon \\
	S &\to aSb
\end{align*}
What this means is that beginning with the string $S$, we keep performing replacements using one of the two rules above until our current string is composed of only terminals.\\
Another example is that of the set of non-empty strings with matched parentheses:
\begin{align*}
	S &\to () \\
	S &\to (S) \\
	S &\to SS
\end{align*}

A context-free grammar is a tuple $(V,\Sigma,R,S)$, where $V$ is a set of \emph{non-terminals} or \emph{variables}, $\Sigma$ is the alphabet, $R : V \to (V \cup \Sigma)^*$ is the finite set of \emph{rules}, and $S \in V$ is the \emph{start symbol}. The language of this grammar is the set of all strings (over terminals) derivable using the rules.\\

Finally, let us look at unrestricted grammars. Consider the language $\{a^nb^nc^n : n \ge 1\}$. This can be represented by the unrestricted grammar
\begin{align*}
	S &\to abc \\
	S &\to aAbc \\
	Ab &\to bA \\
	Ac &\to Bbcc \\
	bB &\to Bb \\
	aB &\to aa \\
	aB &\to aaA
\end{align*}
The only difference between unrestricted and context-free grammars is that the former allow the left-hand side of the rules to be strings as well. That is, an unrestricted grammar is a tuple $(V,\Sigma,R,S)$, where $V$ is a set of non-terminals, $\Sigma$ is the alphabet, $R : (V \cup \Sigma)^* \to (V \cup \Sigma)^*$ is the finite set of rules, and $S \in V$ is the start symbol. As before, the language of this grammar is the set of all strings (over terminals) derivable using the rules.\\

%% LEC 3

Similar to how context-free grammars are equivalent to pushdown automata, \emph{regular expressions} are equivalent to ordinary finite state automata. These are defined as follows:
\begin{enumerate}
	\item $\emptyset$, $\{\epsilon\}$, and $\{a\}$ (for any $a\in\Sigma$) are regular expressions. The second and third sets are often represented as just $\epsilon$ or $a$.
	\item if $E_1, E_2$ are regular expressions,
	\begin{enumerate}
		\item $E_1 + E_2$, where the $+$ represents union, is a regular expression.
		\item $E_1 E_2$, the concatenation of the two expressions, is a regular expression.
		\[ E_1 E_2 = \{ uv : u \in E_1, v \in E_2 \}. \]
		\item $E_1^*$ is a regular expression, where $^*$ is the \emph{Kleene star} operation defined by
		\[ E_1^* = \{u_1u_2\cdots u_n : n \ge 0, u_i \in E_1\text{ for each }i \}. \]
		\item $(E_1)$ is a regular expression.
	\end{enumerate}
\end{enumerate}

So, for example, the language $L$ comprised of the set of all strings with an even number of $a$s over the alphabet $\Sigma = \{a,b\}$ can be defined using the regular expression $b^*(ab^*ab^*)^*$.
As a slightly harder example, the language $L$ comprised of the set of all strings with an even number of $a$s and odd number of $b$s over the alphabet $\Sigma = \{a,b\}$ can be defined as.\\

In pushdown automata, we have a normal FSM along with a stack. The stack has a stack alphabet (that need not be the same as the actual alphabet of the automaton). On each letter of input, we can move to another state and either push a particular symbol onto the stack or pop the topmost symbol. This is usually drawn in the usual FSM way by changing the label of each arrow from just something like $a$ (the read letter) to $a, X | aX$ or $b, aX | X$, which represents that we transition to another state and also push $a$ onto the stack/pop $a$ from the stack. The stack begins with a single symbol $Z_0$, which represents the bottom of the stack. Note that this means that $Z_0$ is always a part of the stack alphabet. Finally, we accept the input iff the stack is empty.