\section{Pre-Requisites}

We assume that the reader has a decent knowledge of General Topology, Analysis, and some parts of Basic Algebra.

Nevertheless, we restate some important topological definitions and results.

\begin{fdef}
	A topological space $X$ is said to be a \textit{Hausdorff space} if for any distinct $x_1,x_2\in X$, there exist disjoint open subsets $U_1$ and $U_2$ such that $x_1\in U_1$ and $x_2\in U_2$.
\end{fdef}

\begin{definition}
	If $X$ is a topological space and $x\in X$, a collection $\mathcal{B}_x$ of neighbourhoods of $x$ is called a \textit{neighbourhood basis for $X$ at $x$} if any neighbourhood of $x$ contains some $B\in\mathcal{B}_x$.\\
	$X$ is said to be \textit{first countable} if there is a countable neighbourhood basis at each point.
\end{definition}

It is often more useful to have a nested neighbourhood basis, defined as follows.

\begin{definition}
	If $X$ is a topological space and $x\in X$, a sequence $(U_n)$ of neighbourhoods of $x$ is said to be a \textit{nested neighbourhood basis at $x$} if $U_{n+1}\subseteq U_n$ for each $n$ and any neighbourhood of $x$ contains some $U_n$.
\end{definition}

The following lemma makes first countable spaces slightly easier to handle.

\begin{lemma}[Nested Neighbourhood Basis Lemma]
	A space $X$ is first countable iff every $x\in X$ has a nested neighbourhood basis.
\end{lemma}

\begin{theorem}
	Let $X$ be a first countable space, $A\subseteq X$, and $x\in X$. Then
	\begin{enumerate}[(a)]
		\item $x\in \overline{A}$ iff $x$ is a limit of a sequence of points in $A$.
		\item $x\in A^\circ$ iff every sequence in $X$ converging to $x$ is eventually in $A$.
		\item $A$ is closed iff it contains every limit of every convergent sequence of points in $A$.
		\item $A$ is open iff every sequence of points converging to a point in $A$ is eventually in $A$.
	\end{enumerate}
\end{theorem}


\begin{fdef}
	A topological space $X$ is said to be \textit{second countable} if it has a countable basis.
\end{fdef}

\begin{theorem}
	\label{theo: second countable then}
	Let $X$ be a second countable space. Then
	\begin{enumerate}[(a)]
		\item $X$ is first countable.
		\item $X$ contains a countable dense subset (it is separable).
		\item Every open cover of $X$ contains a countable subcover.
	\end{enumerate}
\end{theorem}